---
title: "R Colors"
author: "Austin Pesina"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_float: yes
lang: en
---

<!--
Edit this markdown script to change the author: field above and choose a 
different ColorBrewer palette in the Change Color Palette section below.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Introduction

The default color palettes used in computer graphics often leave much to to be desired.  Besides
being aesthetically vapid, they often have poor contrast and colors that are difficult to distinguish,
especially for the color-blind.

Fortunately, cartographic researchers identified this problem quite some time ago.  One of the results
of this research was the ColorBrewer set of palettes.   And lucky for us, these palettes are available
to R programmers in the `rColorBrewer` package.

This demonstration script compares the default palette and a selected ColorBrewer palette.

# Data

Plot some data from a regression example given in Ott and Longneckers _Statistical Methods and Data Analysis_:

```{r read-data}
FER.data <- read.table(header=TRUE, text="
additive	FER
0	1.30
0	1.35
0	1.44
0	1.52
0	1.56
0	1.61
0	1.48
0	1.56
0	1.45
0	1.14
20	2.17
20	2.11
20	2.08
20	2.13
20	2.22
20	2.29
20	2.33
20	2.24
20	2.16
20	2.21
40	2.30
40	2.34
40	2.20
40	2.38
40	2.48
40	2.44
40	2.37
40	2.43
40	2.37
40	2.41
60	2.47
60	2.51
60	2.79
60	2.40
60	2.55
60	2.67
60	2.50
60	2.55
60	2.60
60	2.49
80	3.31
80	3.17
80	3.24
80	3.21
80	3.35
80	3.38
80	3.42
80	3.36
80	3.25
80	3.51
100	4.92
100	3.87
100	4.81
100	4.88
100	5.06
100	5.09
100	4.97
100	4.95
100	4.59
100	4.76
")
```

# Exploratory Plot

Use `tidyverse::ggplot` to draw a _loess_ plot of the data.  Use a color from the 
default color palette. (Colors can be identified by a non-zero integer.)

```{r loess-plot}
ggplot(data=FER.data, aes(x=additive, y=FER)) +
  geom_point() +
  geom_smooth(method="loess", aes(color=3) ) +
  labs(title="Does-Response Experiment") +
  theme_minimal()
```

# Some Fitted Models

Plot a more complicated graph, showing various polynomial fits to the data overlaid on
one another.  Here the default colors are identified by character versions of the integers to 
support the`scale_color_identity_name()` function, used to build a legend for the plot.

```{r fit-models}
ggplot(data=FER.data, aes(x=additive, y=FER)) +
  geom_point() +
  geom_smooth(method="loess", show.legend=TRUE, aes(color="1")) +
  geom_smooth(method="lm", formula='y~x', se=FALSE, show.legend=TRUE, aes(color="2") ) +
  geom_smooth(method="lm", formula='y~poly(x,2)', se=FALSE, show.legend=TRUE, aes(color="3") ) +  
  geom_smooth(method="lm", formula='y~poly(x,3)', se=FALSE, show.legend=TRUE, aes(color="4") ) +  
  scale_color_identity(name = "Model fit",
                          breaks = c("1", "2", "3", "4"),
                          labels = c("Loess", "Linear", "Quadratic", "Cubic"),
                          guide = "legend") +
  labs(title="Chicken Feed Additive", subtitle="polynomial fits") +
  theme_minimal()
```

# Change Color Palette

Hello, ColorBrewer!

```{r}
library("RColorBrewer")
display.brewer.all(colorblindFriendly=TRUE)
```

Redraw the previous graphic, using your favorite ColorBrewer palette:

```{r}
cbPalette <- brewer.pal(5, "YlOrRd")
ggplot(data=FER.data, aes(x=additive, y=FER)) +
  geom_point() +
  geom_smooth(method="loess", show.legend=TRUE, aes(color=cbPalette[1])) +
  geom_smooth(method="lm", formula='y~x', se=FALSE, show.legend=TRUE, aes(color=cbPalette[2]) ) +
  geom_smooth(method="lm", formula='y~poly(x,2)', se=FALSE, show.legend=TRUE, aes(color=cbPalette[3]) ) +  
  geom_smooth(method="lm", formula='y~poly(x,3)', se=FALSE, show.legend=TRUE, aes(color=cbPalette[4]) ) +  
  scale_color_identity(name = "Model fit",
                          breaks = cbPalette[1:4],
                          labels = c("Loess", "Linear", "Quadratic", "Cubic"),
                          guide = "legend") +
  labs(title="Chicken Feed Additive", subtitle="polynomial fits") +
  theme_minimal()
```
